---
title: "TCGA analysis"
author:
- name: Romana Pop
  affiliation: Centre for Molecular Medicine Norway
  email: romanatp@uio.no
package: MARMOT
output:
  BiocStyle::html_document
abstract: |
  Description of your vignette
vignette: |
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

##########################################
# notes to self
* the methylation in ovarian cancer has "-" as separators in the sample IDs
* changed it manually for now, but maybe make some code at some point to go through all the data and make sure the samples names are the same.
* for some reason, feature names seem not to be a thing when loading the data and creating the omic list. Gotta look into that too
###########################################

# Introduction

Some text. 

# Loading libraries and data
```{r}
rm(list=ls())
# load libraries
library(devtools)
devtools::load_all("/storage/kuijjerarea/romana/TCGAmofa/MOFAnet/")
library(tidyverse)
library(preprocessCore)
library(RColorBrewer)
library(msigdbr)
```

# Preparing data
```{r params}
# setting working directory
wd <- "/storage/kuijjerarea/romana/TCGAmofa/june_2024_run"
#wd <- "/storage/kuijjerarea/romana/TCGAmofa/june_2024_run/gep_liver"
setwd(wd)

# specify data directory
data_dir <- "/storage/kuijjerarea/romana/TCGAmofa/cantini_benchmark/momix-notebook/data/cancer/"
#data_dir <- "/storage/kuijjerarea/romana/TCGAmofa/june_2024_run/liver_val_old/"

# specify directory with clinical data
clin_dir <- "/storage/kuijjerarea/romana/TCGAmofa/cantini_benchmark/momix-notebook/data/clinical/"
#clin_dir <- data_dir

# defining vector of cancer names for which to do the analysis
cancers <- c("aml", "colon", "gbm", "kidney", "liver", "lung",
            "melanoma", "ovarian", "sarcoma")
#cancers <- "liver"

model <- c("nonet", "indeg", "out", "both")

# define vector of omic names that will be used
omic_names <- c("expression", "methylation", "miRNA", "indegree", "outdegree")
#omic_names <- c("expression", "indegree", "outdegree")
```
```{r prepare_data}
# prepare data for JDR
for (cancer in cancers) {
    print(cancer)
    # get omic file names
#    files <- c(paste0(rep(paste0(data_dir, cancer, "/"), 5),
#               c("log_exp", "methy", "log_mirna",
#               "indegree_quant.RData", "outdegree.RData")))
    files <- c(paste0(rep(paste0(data_dir, "/"), 3),
               c("exp_log.txt", "indegree_quant.RData", "outdegree.RData")))

    # normalise indegrees
    if (!file.exists(files[4])) {
        #load(paste0(data_dir, cancer, "/indegree.RData"))
        load(paste0(data_dir, "/indegree.RData"))
        indegree <- normalize.quantiles(as.matrix(indegree), copy = FALSE)
        #save(indegree, file = paste0(data_dir, cancer, "/indegree_quant.RData"))
        save(indegree, file = paste0(data_dir, "/indegree_quant.RData"))

    }

    # without PCA
    omics <- prepare_data(omics = files, names = omic_names, pca = FALSE,
                         logs = TRUE,
                         log_name = "logs/prep_data_no_pca_log.txt")
    save(omics, file = paste0(wd, "/", cancer, "_omics_no_pca.Rda"))

    # with PCA
    omics <- prepare_data(omics = files, names = omic_names, pca = TRUE,
                         logs = TRUE, log_name = "logs/prep_data_pca_log.txt",
                         file_name = paste0(cancer, "_omics_pca_results.Rda"))
    save(omics, file = paste0(wd, "/", cancer, "_omics_pca.Rda"))
}
```

```{r prepare_surv}
for (cancer in cancers) {
  print(cancer)
  #clin <- paste0(clin_dir, cancer)
  clin <- paste0(clin_dir, cancer)
  
  # kidney just gotta be special
  if (cancer == "kidney") {
    surv <- prepare_surv(clinical = clin, feature_names = list(sample_id = "submitter_id.samples", 
                                              vital_status = "vital_status.diagnoses",
                                              time_to_event = c("days_to_death.diagnoses",
                                                                "days_to_last_follow_up.diagnoses")))
    surv$sample_id <- str_sub(surv$sample_id, end = -2)
  } else {
    #surv <- prepare_surv(clinical = clin, feature_names = list(sample_id = "sampleID", 
    #                                          vital_status = "vital_status",
    #                                          time_to_event = c("days_to_death", "days_to_last_followup")))
    surv <- prepare_surv(clinical = clin, feature_names = list(sample_id = "Run",
                                              vital_status = "OS.event",
                                              time_to_event = c("OS")))  
  }

  surv$sample_id <- gsub("-", "\\.", surv$sample_id)
  save(surv, file = paste0(wd, "/", cancer, "_surv.Rda"))
}
```

# PCA vs no PCA comparison with multiple JDR tools

```{r bench_no_pca}
for (cancer in cancers) {
    # for now I'm gonna do it with cantini's code until I figure
    # out the finer details for my own code
    source("/storage/kuijjerarea/romana/TCGAmofa/cantini_benchmark/momix-notebook/scripts/runfactorization.R")

    # load data
    load(paste0(wd, "/", cancer, "_omics_no_pca.Rda"))

    factorizations <- runfactorization(omics=omics,num.factors = 5, sep = sep)
    save(factorizations, file = paste0(wd, "/benchmark/", cancer, "_factorisation_no_PCA_5fct.Rda"))

    # run the factorisation with all tools without PCA
    #fct_list <- run_factorisations(omic_list = omics, pca = FALSE, seed = 13)

    #save(fct_list, file = paste0("benchmark/", cancer,
    #     "_JDR_models_no_pca.Rda"))
}
```

```{r bench_pca}
for (cancer in cancers) {
     # for now I'm gonna do it with cantini's code until I figure
    # out the finer details for my own code
    source("/storage/kuijjerarea/romana/TCGAmofa/cantini_benchmark/momix-notebook/scripts/runfactorization.R")

    # load data
    load(paste0(wd, "/", cancer, "_omics_pca.Rda"))

    factorizations <- runfactorization(omics=omics,num.factors = 5, sep = sep)
    save(factorizations, file = paste0(wd, "/benchmark/", cancer, "_factorisation_PCA_5fct.Rda"))
  
    # run the factorisation with all tools without PCA
  #  model_list <- run_factorisations(omic_list = omics, pca = FALSE, seed = 13)

   # save(model_list, file = paste0("benchmark/", cancer,
   #      "_JDR_models_pca.Rda"))

    # extract the factors
  #  for (model in seq_along(model_list)) {
  #      tool <- names(model_list)[model]
  #  }
}
```

# surv analysis for all cancers

```{r surv_assoc_bench}
# with old code for now
#run survival for each cancer for each method
surv_df <- data.frame()
for(i in cancers){
  PCA <- get(load(paste0(wd, "/benchmark/", i,"_factorisation_PCA_5fct.Rda")))
  noPCA <- get(load(paste0(wd, "/benchmark/", i,"_factorisation_no_PCA_5fct.Rda")))
  surv <- get(load(paste0(wd, "/", i, "_surv.Rda"))) # again, this requires the MOFA analysis to have been run first, but for now it works
  
  names(PCA$factorizations) <- PCA$method
  names(noPCA$factorizations) <- noPCA$method
  
  #if(is.null(methods)){
    methods <- PCA$method
  #}
  
  for(j in methods){
    if(j =="MOFA"){
      #get the factorisations for that method
      pca_fct <- PCA$factorizations[[j]][[1]][[1]]
      nopca_fct <- noPCA$factorizations[[j]][[1]][[1]]
    }else{
      #get the factorisations for that method
      pca_fct <- PCA$factorizations[[j]][[1]]
      nopca_fct <- noPCA$factorizations[[j]][[1]]
    }
    
    #run the surv association of the factors
    PCA_cox <- surv_association(pca_fct, surv, univariate=TRUE)
    noPCA_cox <- surv_association(nopca_fct, surv, univariate=TRUE)
    
    df <- surv_compare(models = list(PCA_cox, noPCA_cox), 
                       model_labels = c("PCA", "no_PCA"), univariate = TRUE, 
                      method = "BH")
    
    df$cancer <- i
    df$method <- j
    
    surv_df <- rbind(surv_df, df)
    
    cox_all <- list(PCA_cox, noPCA_cox)
    names(cox_all) <- c("PCA", "no_PCA")
    
    save(cox_all, file=paste0(i,"_cox_models_", j,"_5fct.Rda"))
  }
}
save(surv_df, file=paste0(wd, "/benchmark/TCGA_surv_all_bench_5fct.Rda"))

# extract the factors

```

```{r surv_compare_bench}
load("TCGA_MOFA_surv_all_bench_5fct.Rda")

method <- unique(surv_df$method)
cols <- palette("Dark2")

for(i in method){
  surv_meth <- surv_df[which(surv_df$method == i),]
  models <- unique(surv_meth$label)
  p <- surv_compare_dotplot(surv_df = surv_meth, models_to_compare = models,
                            colours = c(cols[1], "grey", "turquoise"))
  
  ggsave(p, file=paste0(wd,"/figures/surv_compare_PCA_",i,"_5fct.png"))
  ggsave(p, file=paste0(wd,"/figures/surv_compare_PCA_",i,"_5fct.pdf"))
}
```

```{r mofa_models}
for (cancer in cancers) {
  load(paste0(wd, "/", cancer, "_omics_pca.Rda"))

  #nonet
  #data_nonet <- omics[1:3]
  data_nonet <- omics[1]
  mofa_nonet <- run_mofa2(data_nonet, n_fct = 5, seed = 13, convergence = "slow", use_basilisk = T)
  save(mofa_nonet, file = paste0("MOFA_", cancer, "_pca_nonet.Rda"))

  # with indeg
  #data_indeg <- omics[-5] #tcga
  data_indeg <- omics[-3]
  mofa_indeg <- run_mofa2(data_indeg, n_fct = 5, seed = 13, convergence = "slow", use_basilisk = T)
  save(mofa_indeg, file = paste0("MOFA_", cancer, "_pca_indeg.Rda"))
  
  # with outdeg
  #data_out <- omics[-4]
  data_out <- omics[-2]
  mofa_out <- run_mofa2(data_out, n_fct = 5, seed = 13, convergence = "slow", use_basilisk = T)
  save(mofa_out, file = paste0("MOFA_", cancer, "_pca_out.Rda"))
  
  # with both
  data_both <- omics
  mofa_both <- run_mofa2(data_both, n_fct = 5, seed = 13, convergence = "slow", use_basilisk = T)
  save(mofa_both, file = paste0("MOFA_", cancer, "_pca_both.Rda"))
}
```

```{r var_explained_heat}
for (cancer in cancers) {
  load(paste0("MOFA_", cancer, "_pca_nonet.Rda"))
  load(paste0("MOFA_", cancer, "_pca_indeg.Rda"))
  load(paste0("MOFA_", cancer, "_pca_out.Rda"))
  load(paste0("MOFA_", cancer, "_pca_both.Rda"))

  pdf(file = paste0("figures/", cancer, "_var_heat.pdf"))
  n <- plot_variance_explained(mofa_nonet)
  #n + title(main = "no net")
  print(n)
  i <- plot_variance_explained(mofa_indeg)
  #i + title(main = "indegree")
  print(i)
  o <- plot_variance_explained(mofa_out)
 # o + title(main = "outdegree")
  print(o)
  b <- plot_variance_explained(mofa_both)
 # b + title(main = "both")
  print(b)
  dev.off()
}

```

```{r surv_assoc}
# old code
surv_df <- data.frame()
for (i in cancers) {
  # load mofa models
  nonet <- get(load(paste0("MOFA_", i, "_pca_nonet.Rda")))
  deg <- get(load(paste0("MOFA_", i, "_pca_indeg.Rda")))
  out <- get(load(paste0("MOFA_", i, "_pca_out.Rda")))
  both <- get(load(paste0("MOFA_", i, "_pca_both.Rda")))
  
  # get surv data
  if(file.exists(paste0(i,"_surv.Rda"))){ #just for the 10fct thing. make sure to get rid of the "../" later
    load(paste0(i, "_surv.Rda"))
  }else{
    surv <- prepare_surv(clinical = paste0(clin_dir,i),
                         feature_names = list(sample_id = "sampleID", 
                                              vital_status = "vital_status",
                                              time_to_event = c("days_to_death", "days_to_last_followup")))
    surv$sample_id <- gsub("-", "\\.", surv$sample_id)
    
    
    save(surv, file=paste0(i,"_surv.Rda"))
  }
  
  # getting factors
  nonet_fct <- get_factors(nonet)[[1]]
  deg_fct <- get_factors(deg)[[1]]
  out_fct <- get_factors(out)[[1]]
  both_fct <- get_factors(both)[[1]]
  
  # survival association
  # cox <- calculate_surv_association(nonet_fct, surv, by_factor = by_factor)
  # cox_d <- calculate_surv_association(deg_fct, surv, by_factor = by_factor)
  # cox_o <- calculate_surv_association(out_fct, surv, by_factor = by_factor)
  # cox_b <- calculate_surv_association(both_fct, surv, by_factor = by_factor)
  
  cox <- surv_association(nonet_fct, surv, univariate = TRUE)
  cox_d <- surv_association(deg_fct, surv, univariate = TRUE)
  cox_o <- surv_association(out_fct, surv, univariate = TRUE)
  cox_b <- surv_association(both_fct, surv, univariate = TRUE)
  
  # df <- surv_compare(models = list(cox, cox_d, cox_o, cox_b), 
  #                    model_labels = model, by_factor = by_factor, 
  #                    logtrans = logtrans, p_adjust = p_adjust, method = method)
  
  df <- surv_compare(models = list(cox, cox_d, cox_o, cox_b), 
                     model_labels = model, univariate = TRUE, 
                     method = method)
  
  df$cancer <- i
  
  surv_df <- rbind(surv_df, df)
  
  cox_all <- list(cox, cox_d, cox_o, cox_b)
  names(cox_all) <- c("nonet", "indeg", "out", "both")
  
  save(cox_all, file=paste0(i,"_cox_models_PCA_quant.Rda"))
  
  
}

save(surv_df, file=paste0("TCGA_MOFA_surv_df_all_PCA_quant.Rda")) 
```

```{r surv_compare_plots}
# generate comparison plots
if(!exists("surv_df")){
  load(paste0("TCGA_MOFA_surv_df_all_PCA_quant.Rda"))
}

#get models to compare
model_comp <- setdiff(model, "nonet")

for(i in model_comp){
  models_to_compare <- c("nonet", i)
  p <- surv_compare_dotplot(surv_df = surv_df, models_to_compare = models_to_compare)
  
  ggsave(p, file=paste0(wd, "/figures/surv_compare_",models_to_compare[2],"_bee_5_PCA_quant.png"))
  ggsave(p, file=paste0(wd, "/figures/surv_compare_",models_to_compare[2],"_bee_5_PCA_quant.pdf"))
}
```

# liver analysis

```{r clin_association_liver}
clin <- read.table(paste0(clin_dir, "liver"), head = TRUE, sep = "\t")
#clin$sampleID <- gsub("-", "\\.", clin$sampleID)

# get mofa model
#MOFAmodel <- get(load(paste0("MOFA_liver_pca_indeg_quant.Rda")))
MOFAmodel <- get(load(paste0("MOFA_liver_pca_indeg.Rda")))


# make sure samples are the same
clin_fil <- clin[clin[, 1] %in% samples_metadata(MOFAmodel)$sample, ]
colnames(clin_fil)[1] <- "sample"

#make sure missing values are NAs
clin_fil <- mutate_all(clin_fil, ~ ifelse(. == "", NA, .))

# extract factors
Z <- as.data.frame(get_factors(MOFAmodel)[[1]])
feat <- c("Phenotype", "Age", "Sex", "Fibrosis", "Tumor.stage", "Tumor.size")
association <- clin_associaton(Z, clin = clin_fil, clin_feat = feat, sample_label = "sample")

save(association, file = paste0("liver_clin_feat_ass_indeg.Rda"))

# plot clin association
load("liver_clin_feat_ass_indeg.Rda")
pdf(file="figures/liver_clin_ass_indeg_values.pdf")
p <- plot_clin_association(association)
print(p)
dev.off()
```

# differential indegree analysis

```{r diff_indeg_liver}
# load data
indeg <- get(load("liver_omics_no_pca.Rda"))[["indegree"]]
clin <- read.table(paste0(clin_dir, "liver"), sep = "\t", head = TRUE)
colnames(clin)[1] <- "sample_id"
clin$sample_id <- gsub("-", "\\.", clin$sample_id)
load("liver_surv.Rda")
load("TCGA_MOFA_surv_df_all_PCA_quant.Rda")
load("MOFA_liver_pca_both.Rda")
minprops <- c(0.1, 0.2, 0.3)

# get significant factors
sig <- surv_df %>% filter(cancer == "liver", label == "both", padj <= 0.05)
Z <- get_factors(mofa_both)[[1]]
Z <- as.data.frame(Z[, sig$factor])

# run with limma and no covariate correction

# Create the  file names
base_names <- paste0("diff_indeg_liver_limma_", colnames(Z))

# Apply differential_analysis to each column of Z with corresponding file name
#diff <- Map(function(x, file_name) {
#  factor <- setNames(x, rownames(Z))
#  differential_analysis(omic = indeg, factor = factor, surv = surv, clin = clin,
#  file_name = file_name, covariates = c("age_at_initial_pathologic_diagnosis",
#                                        "gender"), sample_label = "sample_id")
#}, Z, file_names)

#diff <- Map(function(x, file_name) {
#  factor <- setNames(x, rownames(Z))
#  differential_analysis(omic = indeg, factor = factor, surv = surv,
#  file_name = file_name, sample_label = "sample_id")
#}, Z, file_names)

# trying this with varying minprop
diff <- Map(function(x, base_name) {
  Map(function(mp) {
    factor <- setNames(x, rownames(Z))
    file_name <- paste0(base_name, "_minprop_", mp, ".Rda")
    differential_analysis(omic = indeg, factor = factor, surv = surv, clin = clin,
      file_name = file_name, covariates = c("age_at_initial_pathologic_diagnosis", "gender"),
      sample_label = "sample_id", minprop = mp)
  }, minprops)
}, Z, base_names)

save(diff, file = "diff_indeg_liver_limma_all_fct_minprop_var.Rda")
```

```{r diff_indeg_liver_gsea}
load("diff_indeg_liver_limma_all_fct_minprop_var.Rda")
minprops <- c(0.1, 0.2, 0.3)

# get pathways
#hallmark <- msigdbr(species = "human", category = "H")
#hallmark <- as.data.frame(cbind(hallmark$gs_name,hallmark$human_gene_symbol))
#colnames(hallmark) <- c("gs_name", "human_gene_symbol")
react <- msigdbr(species = "human", category = "C2", subcategory = "CP:REACTOME")
react <- as.data.frame(cbind(react$gs_name, react$human_gene_symbol))
colnames(react) <- c("gs_name", "human_gene_symbol")

# make pathway names lowercase
#hallmark$gs_name <- tolower(hallmark$gs_name)
react$gs_name <- tolower(react$gs_name)

# remove the "hallmark" before every path name
#hallmark$gs_name <- gsub("hallmark_","",hallmark$gs_name)
react$gs_name <- gsub("reactome_","",react$gs_name)

#hallmark <- split(hallmark$human_gene_symbol, hallmark$gs_name)
react <- split(react$human_gene_symbol, react$gs_name)

base_names <- paste0(rep("liver_gsea_diff_indeg_", length(diff)), names(diff))

# run gsea
#gsea_res <- Map(function(x, file_name) {
#  perform_gsea(diff_results = x, gene_set = hallmark, file_name = file_name,
#  limma = TRUE)}, diff, file_names)

# trying with minprop variation
gsea_res <- Map(function(diff_sublist, base_name) {
  Map(function(x, mp) {
    file_name <- paste0(base_name, "_minprop_", mp, "_react.Rda")
    perform_gsea(diff_results = x, gene_set = react, file_name = file_name, limma = TRUE)
  }, diff_sublist, minprops)
}, diff, base_names)

save(gsea_res, file = "liver_gsea_diff_all_fct_minprop_var_react.Rda")
```

```{r path_select}
load("liver_gsea_diff_all_fct_minprop_var_react.Rda")

# select only pathways that are enriched across all 3 minprops
stable_path <- Map(function(res) {
  select_stable_path(res, thresh = 0.05)
}, gsea_res)


save(stable_path, file = "gsea_stable_path_across_minprop_react.Rda")
```

```{r plot_diff_gsea_results}
load("liver_gsea_diff_all_fct_minprop_var_react.Rda")
load("gsea_stable_path_across_minprop_react.Rda")

gsea_res <- lapply(gsea_res, function(x) x[[1]]) # only for the reactome ones cos I was too lazy to redo a non minprop one separately
titles <- names(gsea_res)
file_names <- paste0(rep("figures/liver_gsea_diff_indeg_", length(gsea_res)), names(gsea_res), "_stable_path_react.pdf")

#get stable pathways across minprop
path_names <- lapply(stable_path, function(x) x[[1]]$pathway)
gsea_subset <- Map(function(x, path) {
  sub <- intersect(x$pathway, path)
  res <- x[which(x$pathway %in% sub), ]
}, gsea_res, path_names)

plots <- Map(function(x, title, file_name, ...) {
  gsea_dotplots(gsea_results = x, surv_df = surv_df, gene_set = "reactome",
  title = title, file_name = file_name, n_path = NULL, thresh = 1.30103,
  width = 160, height = 180, limitsize = FALSE)
}, gsea_subset, titles, file_names)
```

```{r plot_KM_plots}
load("liver_surv.Rda")
load("MOFA_liver_pca_both.Rda")
load("TCGA_MOFA_surv_df_all_PCA_quant.Rda")

sig <- surv_df %>% filter(cancer == "liver", label == "both", padj <= 0.05)
Z <- get_factors(mofa_both)[[1]]
Z <- as.data.frame(Z[, sig$factor])

# plot km
file_names <- paste0(rep("figures/liver_km_", ncol(Z)), colnames(Z), ".pdf")

plots <- Map(function(factor, file_name) {
  factor <- setNames(factor, rownames(Z))
  surv_factor_km(surv = surv, factor = factor, model_label = "both")
  ggsave(file = file_name)
}, Z, file_names)
```

# outdegree analysis

```{r features_mapping_liver}
# load data
out <- get(load("liver_omics_no_pca.Rda"))[["outdegree"]]
load("MOFA_liver_pca_both.Rda")
load("liver_omics_pca_results.Rda")

# get MOFA weights
mofa_weights <- get_weights(mofa_both)

mapped_wts <- Map(function(x, pca_wts) {
  map_wts(fct_weights = x, pca_weights = pca_wts)
}, mofa_weights, omics_pca)

save(mapped_wts, file = "liver_mapped_feat_wts.Rda")
```

```{r plot_feature_weights}
load("liver_mapped_feat_wts.Rda")

factors <- c("Factor2")
file_names <- paste0(rep("figures/liver_feat_wts_", length(omic_names)), omic_names, "_", factors, "_thresh.pdf")

plots <- Map(function(feat, title, file_name, ...) {
  plot_feat_wts(feat_wts = feat, fct = factors, file_name = file_name, thresh = 0.5,
  width = 7, height = 10, limitsize = FALSE)
}, mapped_wts, omic_names, file_names)

```

# randomising mofa seeds

```{r randomise_mofa_seed}
# set directory
seed_dir <- paste0(wd, "/seed_variations/")
setwd(seed_dir)

# set random seed
seeds <- sample.int(100, 5)

for (seed in seeds) {
  for (cancer in cancers) {
    load(paste0(wd, "/", cancer, "_omics_pca.Rda"))

    #nonet
    data_nonet <- omics[1:3]
    mofa_nonet <- run_mofa2(data_nonet, n_fct = 5, seed = seed, convergence = "slow", use_basilisk = T)
    save(mofa_nonet, file = paste0("MOFA_", cancer, "_pca_nonet_", seed, ".Rda"))

    # with both
    data_both <- omics
    mofa_both <- run_mofa2(data_both, n_fct = 5, seed = seed, convergence = "slow", use_basilisk = T)
    save(mofa_both, file = paste0("MOFA_", cancer, "_pca_both_", seed, ".Rda"))
  }
}

save(seeds, file = "random_seeds.Rda")
```

```{r surv_random_seed}
seed_dir <- paste0(wd, "/seed_variations/")
setwd(seed_dir)
load("random_seeds.Rda")

for (seed in seeds) {
  surv_df <- data.frame()
  for (i in cancers) {
    # load mofa models
    nonet <- get(load(paste0("MOFA_", i, "_pca_nonet_", seed, ".Rda")))
    both <- get(load(paste0("MOFA_", i, "_pca_both_", seed, ".Rda")))

    # get surv data
    load(paste0(wd, "/", i, "_surv.Rda"))

    # getting factors
    nonet_fct <- get_factors(nonet)[[1]]
    both_fct <- get_factors(both)[[1]]

    cox <- surv_association(nonet_fct, surv, univariate = TRUE)
    cox_b <- surv_association(both_fct, surv, univariate = TRUE)

    df <- surv_compare(models = list(cox, cox_b),
                     model_labels = model[c(1, 4)], univariate = TRUE,
                     method = method)

    df$cancer <- i

    surv_df <- rbind(surv_df, df)

    cox_all <- list(cox, cox_b)
    names(cox_all) <- c("nonet", "both")

    save(cox_all, file = paste0(i,"_cox_models_PCA_quant_", seed, ".Rda"))
  }

save(surv_df, file = paste0("TCGA_MOFA_surv_df_all_PCA_quant_", seed, ".Rda"))
}
```

```{r surv_compare_random_seed}
seed_dir <- paste0(wd, "/seed_variations/")
setwd(seed_dir)
load("random_seeds.Rda")

for (seed in seeds) {
  # generate comparison plots
  if (!exists("surv_df")) {
    load(paste0("TCGA_MOFA_surv_df_all_PCA_quant_", seed, ".Rda"))
  }

  models_to_compare <- c("nonet", "both")
  p <- surv_compare_dotplot(surv_df = surv_df, models_to_compare = models_to_compare)

  ggsave(p, file=paste0("figures/surv_compare_",models_to_compare[2],"_bee_5_PCA_quant_", seed, ".png"))
  ggsave(p, file=paste0("figures/surv_compare_",models_to_compare[2],"_bee_5_PCA_quant_", seed, ".pdf"))
  
}
```